
CPGTA - Clinical ProteoGenomics Tumor Analysis

Overview
The CPGTA package is designed for processing and analyzing clinical cancer-related ProteoGenomics data, including multi-omics data download, correlation analysis, differential analysis, immunosubtype classification, and survival analysis. The package supports various types of omics data such as transcriptomics, proteomics, and phosphoproteomics.



## Installation
You can install the latest version of the `CPGTA` package from GitHub using the `devtools` package:

```{r}
# Install devtools if not already installed
install.packages("devtools")

# Install cptca from GitHub
devtools::install_github("LYING7797/CPGTA")
```
Data Required
The necessary data files are stored in two Google Drive folders: data and data1. Before using the CPGTA package, you need to manually download the files from these folders and place them in the CPGTA package's root directory (the directory where the DESCRIPTION file is located).

Steps:
1.Download the files from the following Google Drive links:
https://drive.google.com/drive/folders/1Vl5YwTx1q7tTx9RHdNMqnQptH8SNUqm4?usp=sharing for the data folder.
https://drive.google.com/drive/folders/1sExi5ZjhfP8QwD9tVvBPLfdaI4HmGCMF?usp=sharing for the data1 folder.

2.Place the files you downloaded into the data and data1 folders inside the root directory of the CPGTA package (the directory where the DESCRIPTION file is located).

3.Check the folder structure:
The root directory of the CPGTA package should contain two folders: data and data1.
The data files (cancer-PDC info.csv, etc.) should be placed inside these respective folders.
Once the files are downloaded and placed in the correct directories, you can load the CPGTA package and begin using it.

Manually Downloading Data
If the data was not automatically downloaded or you need to update the data, you can manually trigger the download function by running:

```{r}
install_data()
```

Notes:
The data is hosted on Google Drive and will be automatically downloaded during the installation of the package.
The download process will place the data in the data1/ directory inside your working directory.
Ensure you have an active internet connection for the download to complete successfully.

## Functions

cptod()
This function processes clinical, biospecimen, and other omics data types, performs data retrieval, and merges them for analysis.

Arguments:
cancer.type: The cancer type to analyze.
(valid range: 
 Ovarian Serous Cystadenocarcinoma
 Colon Adenocarcinoma
 Breast Invasive Carcinoma
 Uterine Corpus Endometrial Carcinoma
 Clear Cell Renal Cell Carcinoma
 Lung Adenocarcinoma
 Pediatric Brain Tumors
 Hepatocellular Carcinoma
 Glioblastoma
 Early Onset Gastric Cancer
 Head and Neck Squamous Cell Carcinoma
 Lung Squamous Cell Carcinoma
 Pancreatic Ductal Adenocarcinoma
 Acute Myeloid Leukemia
 OV
 COAD
 BRCA
 UCEC
 CCRCC
 LUAD
 Brain Cancer
 LIHC
 GBM
 Gastric Cancer
 HNSCC
 LUSC
 PDAC
 AML)
 
data.category: The omics data category 
(valid range:
Biospecimen
Clinical data
Copy Number Variation
Transcriptome
Proteome
Phosphoproteome).

PDC.study.identifier: Optional PDC study identifier.
data.type: Optional data type ("Harmonized","Normalized").
sample.type: Optional sample type ("Normal tissue", "Primary tumor").

Example:

```{r}
result1 <- cptod(
  cancer.type = "BRCA",
  data.category = "Clinical data",
  PDC.study.identifier = NULL,
  data.type = NULL,
  sample.type = NULL)


result2 <- cptod(
  cancer.type = "OV",
  data.category = "Biospecimen",
  PDC.study.identifier = "PDC000113",
  data.type = NULL,
  sample.type = NULL)


result3 <- cptod(
  cancer.type = "OV",
  data.category = "Transcriptome",
  PDC.study.identifier = NULL,
  data.type = "Normalized",
  sample.type = "Primary tumor")


result4 <- cptod(
  cancer.type = "OV",
  data.category = "Transcriptome",
  PDC.study.identifier = NULL,
  data.type = "Normalized",
  sample.type = "Primary tumor")

result5 <- cptod(
  cancer.type = "PDAC",
  data.category = "Proteome",
  PDC.study.identifier = NULL,
  data.type = c("Normalized","Harmonized"),
  sample.type = "Primary tumor")

result6 <- cptod(
  cancer.type = "UCEC",
  data.category = "Phosphoproteome",
  PDC.study.identifier = "PDC000125",
  data.type = c("Harmonized","Normalized"),
  sample.type = c("Primary tumor","Normal tissue"))

```


cptca()
This function performs correlation analysis between two omics categories (e.g., Transcriptome and Proteome) for a specified cancer type and gene. It returns correlation data and a boxplot visualization.

Arguments:
gene.name: The name of the gene to focus on.
cancer.type: The cancer type to analyze (e.g., "BRCA").
data.category: A vector containing two omics categories ("Transcriptome", "Proteome", "Phosphoproteome"，e.g., c('Transcriptome', 'Proteome')).

```{r}
res <- cptca(
  gene.name    = "TP53",
  cancer.type  = "OV",
  data.category = c("Transcriptome", "Proteome"))
```

cptda()
This function performs differential expression analysis using log2 fold-change threshold for gene expression data and generates a volcano plot.

Arguments:
@param cancer.type: The cancer type to analyze (e.g., "BRCA").
@param data.category: The omics category ("Transcriptome"， "Proteome").
@param log2FC: The log2 fold-change threshold(Must be numeric, with at most one decimal place.).
@param top.gene: The number of top genes to plot based on significance (maximum of 20).

```{r}
res <- cptda (
  cancer.type  = "BRCA",
  data.category = "Proteome",
  log2FC       = 1.5,  
  top.gene     = 10    
)

# The return value res$diff_table is the differential analysis results table (gene, log2FC, p.value, negLog10P, diff), and res$volcano_plot is a ggplot2 object that can be saved or edited further.

```

cptic()
This function retrieves data related to specific PDC identifiers and classified samples for a given cancer type and omics category.

Arguments:
@param cancer.type: The cancer type (e.g., "BRCA").
@param data.category: The omics data category  "Transcriptome" "Proteome").

example:
```{r}
result_rna <- cptic(cancer.type = "BRCA", data.category = "Transcriptome")
print(result_rna)

result_pro <- cptic(cancer.type = "GBM", data.category = "Proteome")
print(result_pro)

```


cptsa()
This function performs survival analysis based on gene expression data for a given cancer type and gene of interest.

Arguments:
@param cancer.type: The cancer type (e.g., "OV").
@param data.category: The omics data category ("Transcriptome", "Proteome").
@param gene.name: The gene to analyze.

Example:
```{r}
res <- cptsa(
  cancer.type  = "OV",
  data.category = "Proteome",
  gene.name = "TP53"
  )
```

License
This package is licensed under the MIT license. See the LICENSE file for details.

